\ourTitle

### 2022-06-11 paper/manuscript.tex
JPC: Added Lead Paragraph for Chaos.

>>>>

{\bf
\section*{Lead:}
Discovering the hidden structure of complex systems has been a widely-recognized
goal of nonlinear science for decades, originally starting with extracting
``geometry'' from a time series. The present results follow directly in this
long line of inquiry, especially that focusing on the causal states or
effective theories underlying time series generated by complex nonlinear
systems. It provides workable statistical estimation methods and dynamical
interpretations of hidden mechanisms. It establishes a new mathematical
foundation for causal inference of complex systems, partly by making
connections to modern machine learning. In this, its use of Cantor embeddings and Wasserstein distances complements recent work on reproducing-kernel Hilbert space representations.
}
